---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/ValeriVoev/danstat.svg?branch=master)](https://travis-ci.org/ValeriVoev/danstat)
<!-- badges: end -->

## Installation
From Github:
```{r, eval = FALSE}
# install.packages("devtools")
devtools::install_github("ValeriVoev/danstat")
```

## Overview
The `danstat` package provides an R interface to Danmarks Statistik Statistikbank API to enable an easier access to the wealth of data in the data bank for research and the general community. The documentation of the API can be found here: [Statistikbank API](https://www.dst.dk/en/Statistik/statistikbanken/api).

## Usage
The API has 4 endpoints which are mimicked by four main functions of the package:

1. `get_subjects()` (SUBJECTS endpoint) retrieves information about subjects around which the data tables in the data bank are organized. The subjects are arranged hierarchically highest level like "Population and elections", "Living conditions", etc. `get_subjects()` retrieves the highest level of the hierarchy. See the function documentation for more details.
1. `get_tables()` (TABLES endpoint) retrieves a list of tables associated with a given subject code. For example `get_tables(subjects = "02")` retrieves all tables related to the subject "Population and elections" with table id, description, variables in the table, etc. 
1. `get_table_metadata()` (TABLEINFO endpoint) returns information about a particular table - description, time of last update, whether or not it is actively updated, and most importantly (for practical purposes) the variable names and id's which are needed whenever you request actual data from the table. Set `variables_only = TRUE` if you only want to get information on the table variables.
1. `get_data()` (DATA endpoint) - returns data from a selected table. It is required to include a `variables` argument as a list. Each element of the list should itself be a named list (with elements `code` and `values`) where `code` is the variable id for which data is requested, and `values` is a vector of values for this variable. If all values are requested, specify `values = NA`. For example:

```{r, message = FALSE}
library(danstat)
user_input = list(list(code = "ieland", values = c(5100, 5128)),
                  list(code = "køn", values = c(1,2)),
                  list(code = "tid", values = NA))

get_data(table_id = "folk1c", variables = user_input)
```
Note that while default language is set to English and variable values are indeed returned in english, e.g. "Men", column names are returned in Danish, e.g. "KØN", "INDHOLD", etc. Unfortunately, the API doesn't currently provide an option to return column names (variable names) in English. However, you can get the English translation using `get_table_metadata`. For example, for the above table 
```{r, message = FALSE}
library(dplyr)
get_table_metadata(table_id = "folk1c", variables_only = TRUE) %>% 
	select(id, text)
```
we can see that "Område" translates to "region", "Køn" to "sex", "Alder" to "age", etc. "Indhold" is always the "value" column whenever data is returned with the `get_data` function.

## See also
There are (as far as I know) two other packages with similar functionality:

* https://github.com/rOpenGov/dkstat
* https://github.com/mikkelkrogsholm/statsDK

In the packages above, the API is called with a `GET` request, while `POST` is the prefrerred option of the API developers and is also what is used in this package. Also, I think that using `POST` requests makes the package code more readable compared to the long url-encoded queries needed for `GET` requests. Also, as of this moment, the rOpenGov package seems to not have been maintained for the past 3 years. In any case, users can consider the above 2 packages as alternatives to this one. 
